Perturbing SU($n$)-symmetric interactions

### 2020-03-17 collective_sun/sun_perturbation.tex
changed a section title and a bit of text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Operator products restricted to the permutationally symmetric
  manifold}
\label{sec:sym_prod}

Weak $M$-body perturbations $\V_M$ to SU($n$)-symmetric interactions
can be treated perturbatively, resulting in an effective Hamiltonian
on a permutationally symmetric manifold $\M_0$ of $N$ spins.
Computing this effective Hamiltonian generally requires multiplying
several operators, and projecting the resulting product onto $\M_0$.
Even if the perturbation $\V_M$ is too strong for rigorous
perturbative treatment, as long as it is comparable in magnitude to
the energetic penalty incurred by breaking permutational symmetry, its
effects can still be accounted for by an effective theory on the joint
manifold of permutationally symmetric states together with states of
``weak'' permutational symmetry breaking, e.g.~the manifold of states
generated by the action of $M$-body perturbations on permutationally
symmetric states.  The expansion of operators in a basis for this
extended manifold likewise requires multiplying several operators and
projecting the resulting product onto $\M_0$.  Specifically, if
$\V_M=\sum_{k\in\C_N\p{M}}w_kO_k$, then we can define operators
\begin{align}
  \V_M^{(k)} \equiv \sum_{p\in\C_N\p{M}} w^{(k)}_p O_p,
\end{align}
where $\set{w^{(k)}}$ forms a complete basis for the set of symmetric
tensors of dimension $M$ with zeros on all diagonals (i.e.~zero in all
entries with a repeated index).  Denoting a complete basis of
permutationally symmetric states by $\set{\ket{m}}$, we can then
identify the $M$-body perturbed states
\begin{align}
  \ket{mk} \equiv \f1{\sqrt{\N_{mk}}} \V_M^{(k)} \ket{m},
  &&
  \N_{mk} \equiv \bk{m | \V_M^{(k)} \V_M^{(k)} | m}.
\end{align}
The restriction of a multi-body operator $\Q=\sum_p v_p Q_p$ to this
expanded Hilbert space is then
\begin{align}
  \Q =_M \P_M \Q \P_M,
  &&
  \P_M \equiv \sum_m \op{m} + \sum_{m,k} \op{mk},
\end{align}
where $=_M$ denotes equality up to a restriction to the support of
$\P_M$, and finding matrix elements of $\Q$ now requires computing
products of the form
\begin{align}
  \bk{\ell | \V_M^{(j)} \Q \V_M^{(k)} | m}
  = \sum_{p,q,r} w^{(j)}_p v_q w^{(k)}_r \bk{\ell | O_p Q_q O_r | m}.
  \label{eq:example_prod}
\end{align}
The purpose of this section is to work out such products in a
systematic way.  To this end, we derive a general diagrammatic
expansion of these products, as well as rules to simplify the
resulting expansions so as to minimize their computational cost.
These diagrammatic methods allow us to algorithmically evaluate
products such as \eqref{eq:example_prod} from the bare multi-body
operator data $\p{w^{(j)},O},\p{v,Q},\p{w^{(k)},O}$.  We then work out
some special cases of products that are relevant to the perturbation
theory in Sections \ref{sec:single_body_pert} and
\ref{sec:two_body_pert}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Projecting products of operators onto the permutationally
  symmetric manifold}
\label{sec:sym_prod}

Weak $M$-body perturbations $\V_M$ to SU($n$)-symmetric interactions
can be treated perturbatively, resulting in an effective Hamiltonian
on a permutationally symmetric manifold $\M_0$ of $N$ spins.
Computing this effective Hamiltonian generally requires multiplying
several operators, and projecting the resulting product onto $\M_0$.
Even if the perturbation $\V_M$ is too strong for rigorous
perturbative treatment, as long as it is comparable in magnitude to
the energetic penalty incurred by breaking permutational symmetry, its
effects can still be accounted for by an effective theory on the joint
manifold of permutationally symmetric states together with states of
``weak'' permutational symmetry breaking, e.g.~the manifold of states
generated by the action of $M$-body perturbations on permutationally
symmetric states.  The expansion of operators in a basis for this
extended manifold likewise requires multiplying several operators and
projecting the resulting product onto $\M_0$.  Specifically, if
$\V_M=\sum_{k\in\C_N\p{M}}w_kO_k$, then we can define operators
\begin{align}
  \V_M^{(k)} \equiv \sum_{p\in\C_N\p{M}} w^{(k)}_p O_p,
\end{align}
where $\set{w^{(k)}}$ forms a complete basis for the set of symmetric
tensors of dimension $M$ with zeros on all diagonals (i.e.~zero in all
entries with a repeated index).  Denoting a complete basis of
permutationally symmetric states by $\set{\ket{m}}$, we can then
identify the $M$-body perturbed states
\begin{align}
  \ket{mk} \equiv \f1{\sqrt{\N_{mk}}} \V_M^{(k)} \ket{m},
  &&
  \N_{mk} \equiv \bk{m | \V_M^{(k)} \V_M^{(k)} | m}.
\end{align}
The restriction of a multi-body operator $\Q=\sum_p v_p Q_p$ to this
expanded Hilbert space is then
\begin{align}
  \Q =_M \P_M \Q \P_M,
  &&
  \P_M \equiv \sum_m \op{m} + \sum_{m,k} \op{mk},
\end{align}
where $=_M$ denotes equality up to a restriction to the support of
$\P_M$, and finding matrix elements of $\Q$ now requires computing
products of the form
\begin{align}
  \bk{\ell | \V_M^{(j)} \Q \V_M^{(k)} | m}
  = \sum_{p,q,r} w^{(j)}_p v_q w^{(k)}_r \bk{\ell | O_p Q_q O_r | m}.
  \label{eq:example_prod}
\end{align}
The purpose of this section is to work out such products in a
systematic way.  To this end, we derive a general diagrammatic
expansion for these products, as well as a set of rules to simplify
these expansions so as to minimize their computational cost.  These
diagrammatic methods allow us to algorithmically evaluate products
such as \eqref{eq:example_prod} from the bare multi-body operator data
$\p{w^{(j)},O},\p{v,Q},\p{w^{(k)},O}$.  We then work out some special
cases of products that are relevant to the perturbation theory in
Sections \ref{sec:single_body_pert} and \ref{sec:two_body_pert}.
